---
title: "p8105_hw3_ec3663"
author: "Eunsuh Cho"
date: "2023-10-13"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Problem 1

## Read in the data

```{r}
data("instacart")

instacart = 
  instacart |> 
  as_tibble()
```

## Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

# Problem 2

## Load packages and data

```{r p2_setup}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(p8105.datasets)
data("brfss_smart2010")
```

## Clean data

```{r}
brfss_df = brfss_smart2010 |> 
  janitor::clean_names() |> 
  rename(state = locationabbr, county = locationdesc) |> 
  filter(topic == "Overall Health",
    response %in% c("Poor", "Fair", "Good", "Very Good", "Excellent")) |> 
  mutate(response = factor(
    response, levels = c("Poor", "Fair", "Good", "Very Good", "Excellent")))
```

## In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss_02_df = brfss_df |> 
  filter(year == "2002") |> 
  group_by(state) |> 
  summarise(locations = n_distinct(county)) |> 
  filter(locations >= 7)

print(brfss_02_df)

brfss_10_df = brfss_df |> 
  filter(year == "2010") |> 
  group_by(state) |> 
  summarise(locations = n_distinct(county)) |> 
  filter(locations >= 7)

print(brfss_10_df)
```

In 2002, `r with(brfss_02_df, paste(state), collapse = ", ")` were observed at 7 or more locations.

In 2010, `r with(brfss_10_df, paste(state), collapse = ", ")` were observed at 7 or more locations.

## Construct a dataset limited to `Excellent` responses

```{r}
brfss_ex_df = brfss_df |> 
  filter(response == "Excellent") |> 
  group_by(state, year) |> 
  mutate(avg_data_value = mean(data_value, na.rm = TRUE)) |> 
  select(year, state, avg_data_value)
```

The resulting dataset that is limited to `Excellent` responses has `r nrow(brfss_ex_df)` observations.

### Create a "spaghetti" plot of the average value over time within a state

```{r}
ggplot(brfss_ex_df, aes(x = year, y = avg_data_value, group = state, color = state)) +
  geom_line() +
  labs(title = "Average Data Value by State over Years", x = "Year", y = "Average Data Value", color = "State") +
  theme_minimal()
```

The "spaghetti" plot shows a general decrease in average data value over time. The range from 17.5 to 27.5 contains the majority of the average data values, but some fall outside of this range. The state of WV in particular is a visible outlier.

## Make a two-panel plot showing, for 2006 and 2010, the `data_value` distribution for responses ("Poor" to "Excellent") among NY State locations





